Bird beak's elaboration and exploration

### 2021-10-11 Analysis/03.1-results.Rmd
Update 03.1-results.Rmd

```{r, echo = FALSE, fig.height = 18, fig.width = 18}
par(mfrow = c(2,2))
## Plotting mean axes
sauron.plot(superorder_shapespace,
            col = c(rainbow(n.subsets(superorder_shapespace)-1), "grey"),
            main = "Average ellipses by super order",
            ellipses = mean,
            legend = FALSE,
            points = FALSE,
            cex = 0.5)
## Plotting mean ellipses
sauron.plot(superorder_shapespace,
            col = c(rainbow(n.subsets(superorder_shapespace)-1), "grey"),
            main = "Average major axes by super order",
            major.axes = mean,
            legend = TRUE,
            points = FALSE,
            cex = 0.5)
## Plotting mean axes
sauron.plot(order_shapespace,
            col = c(rainbow(n.subsets(order_shapespace)-1), "grey"),
            main = "Average ellipses by order",
            ellipses = mean,
            legend = FALSE,
            points = FALSE,
            cex = 0.5)
## Plotting mean ellipses
sauron.plot(order_shapespace,
            col = c(rainbow(n.subsets(order_shapespace)-1), "grey"),
            main = "Average major axes by order",
            major.axes = mean,
            legend = TRUE,
            points = FALSE,
            cex = 0.5)
```
Main axes and ellipses visualisation

@@

```{r, fig.height = 20, fig.width = 10, echo = FALSE, eval = TRUE}
## Run the analyses between groups
super_results_between <- covar.projections.wrapper(superorder_shapespace,
                                             type = "between",
                                             base = "phylogeny",
                                             dispRity.out = TRUE)
## Run the analyses between groups
order_results_between <- covar.projections.wrapper(order_shapespace,
                                             type = "between",
                                             base = "phylogeny",
                                             dispRity.out = TRUE)
## Run the analyses of groups on the phylogeny
super_results_phylo <- covar.projections.wrapper(superorder_shapespace,
                                            type = "within",
                                            base = "phylogeny",
                                            incl.base = TRUE,
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses of groups on the phylogeny
order_results_phylo <- covar.projections.wrapper(order_shapespace,
                                            type = "within",
                                            base = "phylogeny",
                                            incl.base = TRUE,
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses groups on themselves
super_results_within <- covar.projections.wrapper(superorder_shapespace,
                                            type = "within",
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses groups on themselves
order_results_within <- covar.projections.wrapper(order_shapespace,
                                            type = "within",
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Save all
allbirds_8748t_8Dshapespace_31levels_results_list <- list(
        "super_results_between" = super_results_between,
        "order_results_between" = order_results_between,
        "super_results_phylo"   = super_results_phylo,
        "order_results_phylo"   = order_results_phylo,
        "super_results_within"  = super_results_within,
        "order_results_within"  = order_results_within)
save(allbirds_8748t_8Dshapespace_31levels_results_list, file = "../Data/Processed/allbirds_8748t_8Dshapespace_31levels_results_list.rda")
```

@@

```{r, eval = TRUE, echo = FALSE, eval = FALSE}
library(hrdcde)

#@param results the results list
#@param type which type (position, distance or angle)
#@param cent.tend which central tendency fun (default is median)
#@param use.hdr whether to use kernel density estimate for the outputs (default is TRUE)
#@param prob which quantiles to use (default is c(50, 95, 99))
#@param ... plot options

get.distributions <- function(results, type, cent.tend = median, use.hdr = TRUE, prob = c(50, 95, 99), ...) {

    ## Extracting all the results
    disparity <- lapply(results[[type]]$disparity, function(X, type) return(X$elements))

@@

    ## Calculate and plot the hdrcde
    op_tmp <- par(mfrow = c(ceiling(sqrt(n_groups)), round(sqrt(n_groups))))
    densities <- list()
    for(i in 1:length(centrals)) {
        if(use.hdr) {
            densities[[i]] <- hdr.den(centrals[[i]], prob = prob, main = names(centrals)[i], ...)
        } else {
            hist(centrals[[i]], main = names(centrals)[i], ...)
        }
    }
    par(op_tmp)

    ## Find the extreme species
    get.extremes.hdr <- function(density, central) {
        ## Get the minimums and maximums
        mins <- which(central <= density$hdr[1,1])
        maxs <- which(central >= density$hdr[1,2])
        ## Get their values
        return(c(central[names(mins)], central[names(maxs)]))
    }
    CI.converter <- function(CI) {
        sort(c(50-CI/2, 50+CI/2)/100)
    }
    get.extremes.quantile <- function(central, prob) {
        ## Get the quantiles
        quants <- quantile(central, probs = CI.converter(max(prob)))
        ## Get the minimums and maximums
        mins <- which(central <= quants[1])
        maxs <- which(central >= quants[2])
        ## Get their values
        return(c(central[names(mins)], central[names(maxs)]))
    }

@@

    ## Return the as a table
    for(i in 1:length(output)) {
        output[[i]] <- data.frame("level" = rep(names(output)[i], length(output[[i]])),
                                  "species" = names(output[[i]]),
                                  "score"   = unname(output[[i]]))
    }
    return(output)
}
```

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = FALSE}
get.distributions(results$super_results_between, type = "position")
```

>>>>

```{r, echo = FALSE, fig.height = 18, fig.width = 18, eval = FALSE}
par(mfrow = c(2,2))
## Plotting mean axes
sauron.plot(superorder_shapespace,
            col = c(rainbow(n.subsets(superorder_shapespace)-1), "grey"),
            main = "Average ellipses by super order",
            ellipses = mean,
            legend = FALSE,
            points = FALSE,
            cex = 0.5)
## Plotting mean ellipses
sauron.plot(superorder_shapespace,
            col = c(rainbow(n.subsets(superorder_shapespace)-1), "grey"),
            main = "Average major axes by super order",
            major.axes = mean,
            legend = TRUE,
            points = FALSE,
            cex = 0.5)
## Plotting mean axes
sauron.plot(order_shapespace,
            col = c(rainbow(n.subsets(order_shapespace)-1), "grey"),
            main = "Average ellipses by order",
            ellipses = mean,
            legend = FALSE,
            points = FALSE,
            cex = 0.5)
## Plotting mean ellipses
sauron.plot(order_shapespace,
            col = c(rainbow(n.subsets(order_shapespace)-1), "grey"),
            main = "Average major axes by order",
            major.axes = mean,
            legend = TRUE,
            points = FALSE,
            cex = 0.5)
```
Main axes and ellipses visualisation

@@

```{r, fig.height = 20, fig.width = 10, echo = FALSE, eval = FALSE}
## Run the analyses between groups
super_results_between <- covar.projections.wrapper(superorder_shapespace,
                                             type = "between",
                                             base = "phylogeny",
                                             dispRity.out = TRUE)
## Run the analyses between groups
order_results_between <- covar.projections.wrapper(order_shapespace,
                                             type = "between",
                                             base = "phylogeny",
                                             dispRity.out = TRUE)
## Run the analyses of groups on the phylogeny
super_results_phylo <- covar.projections.wrapper(superorder_shapespace,
                                            type = "within",
                                            base = "phylogeny",
                                            incl.base = TRUE,
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses of groups on the phylogeny
order_results_phylo <- covar.projections.wrapper(order_shapespace,
                                            type = "within",
                                            base = "phylogeny",
                                            incl.base = TRUE,
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses groups on themselves
super_results_within <- covar.projections.wrapper(superorder_shapespace,
                                            type = "within",
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Run the analyses groups on themselves
order_results_within <- covar.projections.wrapper(order_shapespace,
                                            type = "within",
                                            dispRity.out = TRUE,
                                            verbose = TRUE)
## Save all
allbirds_8748t_8Dshapespace_31levels_results_list <- list(
        "super_results_between" = super_results_between,
        "order_results_between" = order_results_between,
        "super_results_phylo"   = super_results_phylo,
        "order_results_phylo"   = order_results_phylo,
        "super_results_within"  = super_results_within,
        "order_results_within"  = order_results_within)
save(allbirds_8748t_8Dshapespace_31levels_results_list, file = "../Data/Processed/allbirds_8748t_8Dshapespace_31levels_results_list.rda")
```

@@

```{r, eval = TRUE, echo = FALSE, eval = TRUE}
library(hdrcde)

#@param res the results list
#@param type which type (position, distance or angle)
#@param cent.tend which central tendency fun (default is median)
#@param use.hdr whether to use kernel density estimate for the outputs (default is TRUE)
#@param prob which quantiles to use (default is c(50, 95, 99))
#@param exclude.phylo remove the phylogeny subset (default = FALSE)
#@param ... plot options

get.distributions <- function(res, type, cent.tend = median, use.hdr = TRUE, prob = c(50, 95, 99), exclude.phylo = FALSE, ...) {

    ## Remove the phylo one (if needed)
    if(exclude.phylo) {
        select <- names(size.subsets(res[[type]]))
        select <- select[-which(select %in% "phylogeny")]
        res[[type]] <- get.subsets(res[[type]], subsets = select)
    }
    ## Extracting all the results
    disparity <- lapply(res[[type]]$disparity, function(X, type) return(X$elements))

@@

    ## Calculate and plot the hdrcde
    op_tmp <- par(mfrow = c(ceiling(sqrt(n_groups)), round(sqrt(n_groups))))
    densities <- list()
    for(i in 1:length(centrals)) {
        if(use.hdr) {
            densities[[i]] <- hdr.den(centrals[[i]], prob = prob, main = names(centrals)[i])#, ...)
        } else {
            hist(centrals[[i]], main = names(centrals)[i], ...)
        }
    }
    par(op_tmp)

    ## Find the extreme species
    get.extremes.hdr <- function(density, central) {
        ## Get the minimums and maximums
        mins <- which(central <= density$hdr[1,1])
        maxs <- which(central >= density$hdr[1, ncol(density$hdr)])
        ## Get their values
        return(c(central[names(mins)], central[names(maxs)]))
    }
    CI.converter <- function(CI) {
        sort(c(50-CI/2, 50+CI/2)/100)
    }
    get.extremes.quantile <- function(central, prob) {
        ## Get the quantiles
        quants <- quantile(central, probs = CI.converter(max(prob)))
        ## Get the minimums and maximums
        mins <- which(central <= quants[1])
        maxs <- which(central >= quants[2])
        ## Get their values
        return(c(central[names(mins)], central[names(maxs)]))
    }

@@

    ## Return the as a table
    for(i in 1:length(output)) {
        if(length(output[[i]]) == 0) {
            output[[i]] <- data.frame("level" = names(output)[i], "species" = NA, "score" = NA)
        } else {
            output[[i]] <- data.frame("level" = rep(names(output)[i], length(output[[i]])),
                                      "species" = names(output[[i]]),
                                      "score"   = unname(output[[i]]))
        }
    }

    return(output)
}
combine.results <- function(res1, res2) {
    ## Combine the results
    res_out <- rbind(res1[,1:2], res2[, 1:2])
    res_out$elaboration <- c(res1[, 3], rep(NA, nrow(res2)))
    res_out$exploration <- c(rep(NA, nrow(res1)), res2[, 3])
    ## Find any species in commons
    commons <- intersect(res1$species, res2$species)
    if(length(commons) > 0 && is.na(commons)) {
        if(is.na(res1$species)) {
            if(is.na(res2$species)) {
                ## Both are NA (keep just the first one)
                res_out <- res_out[-1, , drop = FALSE]
            } else {
                ## Remove res1
                res_out <- res_out[-1, , drop = FALSE]
            }
        } else {
            ## Remove res2
            res_out <- res_out[1:length(res1$species), , drop = FALSE]
        }
    } else {
        ## Recursively remove the ones in common
        while(length(commons) > 0) {
            duplicates <- which(res_out$species == commons[1])
            res_out[duplicates[1], 4] <- res_out[duplicates[2], 4]
            res_out <- res_out[-duplicates[2], ]
            commons <- commons[-1]
        }
    }
    return(res_out)
}
```

### Extreme species per super order (based on phylogeny)

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = TRUE}
super_positions_phylo <- get.distributions(results$super_results_phylo, type = "position")
```
Figure: elaborations distributions per super orders

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = TRUE}
super_distances_phylo <- get.distributions(results$super_results_phylo, type = "distance")
```
Figure: explorations distributions per super orders

```{r, echo = FALSE}
library(knitr)
## Combining the results
super_phylo_table <- do.call(rbind, mapply(combine.results, super_positions_phylo, super_distances_phylo, SIMPLIFY = FALSE))
rownames(super_phylo_table) <- NULL
kable(super_phylo_table, digits = 3)
```

### Extreme species per order (based on phylogeny)

```{r, eval = TRUE, echo = FALSE, fig.height = 24, fig.width = 24, eval = TRUE}
order_positions_phylo <- get.distributions(results$order_results_phylo, type = "position", exclude.phylo = TRUE)
```
Figure: elaborations distributions per order

```{r, eval = TRUE, echo = FALSE, fig.height = 24, fig.width = 24, eval = TRUE}
order_distances_phylo <- get.distributions(results$order_results_phylo, type = "distance", exclude.phylo = TRUE)
```
Figure: explorations distributions per order

```{r, echo = FALSE}
## Combining the results
order_phylo_table <- do.call(rbind, mapply(combine.results, order_positions_phylo, order_distances_phylo, SIMPLIFY = FALSE))
rownames(order_phylo_table) <- NULL
kable(order_phylo_table, digits = 3)
```

### Extreme species per super order (within)

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = TRUE}
super_positions_within <- get.distributions(results$super_results_within, type = "position", exclude.phylo = TRUE)
```
Figure: elaborations distributions within super orders

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = TRUE}
super_distances_within <- get.distributions(results$super_results_within, type = "distance", exclude.phylo = TRUE)
```
Figure: explorations distributions within super orders

```{r, echo = FALSE}
## Combining the results
super_phylo_table_within <- do.call(rbind, mapply(combine.results, super_positions_within, super_distances_within, SIMPLIFY = FALSE))
rownames(super_phylo_table_within) <- NULL
kable(super_phylo_table_within, digits = 3)
```

### Extreme species per order (within)

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = TRUE}
order_positions__within <- get.distributions(results$order_results_within, type = "position", exclude.phylo = TRUE)
```
Figure: elaborations distributions within orders

```{r, eval = TRUE, echo = FALSE, fig.height = 12, fig.width = 12, eval = TRUE}
order_distances_within <- get.distributions(results$order_results_within, type = "distance", exclude.phylo = TRUE)
```
Figure: explorations distributions within orders

```{r, echo = FALSE}
## Combining the results
order_phylo_table_within <- do.call(rbind, mapply(combine.results, order_positions_within, order_distances_within, SIMPLIFY = FALSE))
rownames(order_phylo_table_within) <- NULL
kable(order_phylo_table_within, digits = 3)
```
